---
title: "Reproduciable Research Assignment 2"
author: "Mao-Lin Li"
date: "August 20, 2015"
output: html_document
---

# Synoposis

In this report, we analyze the U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database to evaluate the factors that affecting population health and economic consequences. We use the data from 1950 to the end in November 2011. We take fatalities and injuries as the reference for evaluating population health and calculate crop damage as the ference for mesuring economic consequences.

```{r echo=TRUE}
knitr::opts_chunk$set(cache=TRUE)
# Load dplyr library
library(dplyr)
# Load data
stormData<-read.csv("repdata-data-StormData.csv", stringsAsFactors=FALSE)
# We select the subset that fatalities, injuries, property damage or crop damage large than zero.
targetData<-filter(stormData, FATALITIES>0 | INJURIES > 0 | PROPDMG > 0 | CROPDMG >0)
```

### 1. Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?

* Data Processing

Here we take the number of fatalities and injuries as the reference for evaluating population health.

```{r echo=TRUE}
# Aggregate the sum of fatalities according to type
fatalData<-aggregate(targetData$FATALITIES, list(type=targetData$EVTYPE), sum, na.rm=TRUE)
names(fatalData)<-c("Type", "Fatalities")
# Aggregate the sum of injuries according to type
injureData<-aggregate(targetData$INJURIES, list(typr=targetData$EVTYPE), sum, na.rm=TRUE)
names(injureData)<-c("Type", "Injuries")
# Aggregate the sum of fatalities and injuries according to type.
impactData<-aggregate(targetData$FATALITIES + targetData$INJURIES, list(type=targetData$EVTYPE), sum, na.rm=TRUE)
names(impactData)<-c("Type", "FatalitiesAndInjuries")
# Merge above three datasets into one data frame
healthResult<-merge(fatalData, injureData)
healthResult<-merge(healthResult, impactData)
healthResult<-healthResult[order(healthResult$FatalitiesAndInjuries, decreasing = TRUE),]
head(healthResult)
```


From above result, we know Tornado brings the most harm to population health.

### 2. Across the United States, which types of events have the greatest economic consequences?

* Data Processing

Here we take the number of crop damage as the reference for evaluating population health.

```{r, echo=TRUE}

```

